Single Site, Single Disturbance

### 2020-10-19 SingleSite_SingleDisturb.Rmd
SSSD: separate R model from JAGS forecast

## Todo
-   Move forecast out of JAGS
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA
-   Calibrate state and transition probabilities for a single region
-   Upgrade code to multinomial
-   Add PFT switching
-   Add ensemble weighting so that we can manage ensemble sizes
-   Add "fusion" if a case becomes too rare to worry about
-   Think about whether we need to add LULC state to state variables & how LULC state would be assimilated
-   Test against subseasonal dynamics (e.g. distinguishing phenology from disturbance)
-   investigate whether we need a "jumpback" capacity once disturbance event type and timing is resolved
-

>>>>

## Foreward filter, separate process model

In the previous code, both the forecast and analysis steps were done within JAGS. In this version we're pulling out the Forecast step and running that in R since (i) that's how most models work and (ii) it will allow us to more easily distinguish the Forecast prior and Analysis posteriors

```{r}
AnalysisSSSD <- "
model{
  ## prior
  muN ~ dnorm(mufN,pfN)
  muD ~ dnorm(mufD,pfD)
  D ~ dbern(p)
  ## Analysis
  n <- D*muD + (1-D)*muN
  Y ~ dnorm(n,r)
}"
## Storage
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Dbar = 0
Analysis <- Forecast <- list()
## set IC
nMC = 5000
Xl = nIC         
pl = 1/n.sd^2
Na = rnorm(nMC,Xl,n.sd)
Nf = N*NA

for(t in 1:NT){
  ## Forecast 
  d.prior <- rbinom(nMC,1,p.dist)   ## prior on disturbance status
  sel <- !as.logical(d.prior)
  Nf[!sel] <- rnorm(sum(d.prior),data$mu0,1/sqrt(data$pa0))  ## draw disturbed predictions
  Nf[sel]  <- rnorm(sum(sel),Na[sel] + r*Na[sel]*(1-Na[sel]/K),n.sd) ## draw undisturbed
  Nf[Nf < 0] <- 0 ## Tobit

  ## priors
  priors <- list(p    = mean(d.prior),
                 mufN = mean(Nf[sel]),
                 mufD = mean(Nf[!sel]),
                 pfN  = 1/var(Nf[sel]),
                 pfD  = 1/var(Nf[!sel])
                 )
  Forecast[[t]] <- c(list(d.prior=d.prior,Nf=Nf),priors)

  ## Analysis
  update = c(list(Y=data$Y[t],r=1/obs.sd^2),priors)
  mod <- jags.model(file=textConnection(AnalysisSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

  ## update parameters  
  dat = as.matrix(jdat)
  Xl  = mean(dat[,"n"])
  pl  = 1/var(dat[,"n"])
  CI.fN[,t] = quantile(dat[,"n"],c(0.025,0.25,0.5,0.75,0.975))
  CI.fD[,t] = quantile(dat[,"D"],c(0.025,0.25,0.5,0.75,0.975))
  Dbar[t] = mean(dat[,"D"])

  ## draw new IC
  Na <- sample(dat[,"n"],nMC,replace=TRUE)

```{r}
## organize Analysis state var (n) into a wide format
n.wide <- sapply(Analysis,function(x){x[,"n"]})
colnames(n.wide) <- paste0("t",1:NT)
n.wide <- as.data.frame(n.wide)

## violin plot
n.long %>%
 filter(n >= 0 & n <= 8) %>%
 ggplot() +
 aes(x = "", y = n, group = time) +
 geom_violin(adjust = 0.5, scale = "area", fill = "#0c4c8a") +
 labs(x = "Time", y = "LAI") +
 theme_minimal()
```

## Todo
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA
-   Calibrate state and transition probabilities for a single region
-   Upgrade code to multinomial
-   Add PFT switching
-   Add ensemble weighting so that we can manage ensemble sizes
-   Add "fusion" if a case becomes too rare to worry about
-   Think about whether we need to add LULC state to state variables & how LULC state would be assimilated
-   Test against subseasonal dynamics (e.g. distinguishing phenology from disturbance)
-   investigate whether we need a "jumpback" capacity once disturbance event type and timing is resolved
-
